### 2018-01-05 latex/egsr17-cgf/introduction.tex
started working on paper text

\section{Introduction}

\subsection*{Contributions}
\begin{itemize}
	\item We derive the time-independent real-valued PN-equations.
	\item We present a novel system for generating finite difference stencils for coupled PDE's, directly from a computer algebra representation of mathematical expressions.
	\item We introduce a staggered-grid solver for time-independent PN-equations and use it for rendering participating media.
	\item Finally, we compare higher order PN against Flux-Limited Diffusion.
\end{itemize}

In the next section we will discuss related work and why it motivates our contribution. Section~\ref{sec:pnequations} introduces the real-valued $P_N$-equations, which we solve using our solver described in section~\ref{sec:pnsolver}. We briefly recapitulate variable eddington factor methods and flux-limited diffusion in particular in section~\ref{sec:vef}. Our solvers for $P_N$ and FLD are run on a set of standard problems in section~\ref{sec:results}. We finish with our conclusion in section~\ref{sec:conclusion}.